---
title: "MBS Comtrade Data Notebook"
output: html_notebook
---


Load the libraries.

```{r}
library(data.table)
library(ggplot2)
```

The code below will show the data cleaning and exploration.

First load the data.

# Data Wrangling

```{r}
setwd("C:/Users/Ferhat/Documents/GitHub/WorldBankData")

data = read.csv("MBSComtrade.csv")

data$series_type = NULL
data$table_type = NULL
data$table_type_desc = NULL
data$period_type = NULL
data$trade_flow = NULL
data$currency_type = NULL
data$value_unit = NULL
data$value_type = NULL
data$value_type_desc = NULL
data$base_year = NULL
data$period_number = NULL
data$country_type = NULL
data$estimation_type = NULL
data$estimation_type_desc = NULL


save(data, file = "cleaned_data.rdata")
```

```{r}
data = as.data.table(data)
```

Lets explore some of the data.

```{r}
head(data[country_english_name == "USA"],10)
```
Lets do some plots. But first lets get rid of world regions and only keep countryes so we don't double count. Lets look at the USA first.
```{r}
USA = data[country_english_name == "USA" & partner_country_code > 0 & partner_country_code < 899 ]
head(USA,10)
```

Now lets group  everything by partner export country.
```{r}
USA_group = USA[, .(value.SUM = sum(value)), by=c("partner_country_english_name", "year")]
setnames(USA_group, c("Country","Year", "TotalValue") )
head(USA_group,10)
```

Now lets plot.

# Line chart

```{r}

library(plotly)
p = ggplot(USA_group, aes(x = Year, y=TotalValue))+
  geom_line(aes(colour = Country))+
  geom_point(aes(colour = Country), size=2)+
  theme(legend.position="none")

ggplotly(p)
```
# Pie chart

```{r}
ggplot(USA_group, aes(x=factor(1), fill = factor(Country)))+
  geom_bar(aes(weight = TotalValue), width = 1)+
  coord_polar(theta = "y", start=0)+
  theme(legend.position="none")
```

# Sankey Diagram

```{r  results='asis'}
USA_group_country = USA_group[, .(Sum = sum(TotalValue)), by=Country]
USA_group_country$Source = "USA"
USA_group_country[, Country:=as.character(Country)]

USA_group_country <- USA_group_country %>% 
  select(Country, Source, Sum)
setorder(USA_group_country, Sum)

slice = tail(USA_group_country, 10)

slice <- slice %>% 
  select(Source, Country, Sum)

library(googleVis)
sk1 <- gvisSankey(slice, from="Source", to="Country", weight="Sum")
print(sk1, 'chart')
```
Lets play with leaflet package.

# Leaflet

```{r}
library(leaflet)
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)
m %>% addTiles()
```
Lets get the latitude and longitude of countries.

```{r}
countries = read.csv("countries.csv")
countries = as.data.table(countries)
setnames(countries, old = c("country", "name"), new = c("CountryCode", "Country"))

head(countries,10)
```
Now lets populate USA_group data table the latitude and longitude of the countries.

```{r}
USA_group_joined = merge(USA_group, countries, by = "Country")

na.omit(USA_group_joined)

head(USA_group_joined, 10)
```

```{r}
setorder(USA_group_joined, Country, Year)
leaflet(USA_group_joined[USA_group_joined[,Year == 2010]]) %>% addTiles() %>%
  addCircles(lng = ~longitude, lat = ~latitude, weight = 1,
    radius = ~sqrt(TotalValue), popup = ~paste(Country,"Volume: ",TotalValue), fillOpacity = 0.5
  )
```

Now lets look at the flows using Leaflet's addflows function.

```{r}

leaflet(USA_group_joined[USA_group_joined[,Year == 2010]]) %>% addTiles() %>%
  addCircles(lng = ~longitude, lat = ~latitude, weight = 1,
    radius = ~sqrt(TotalValue), popup = ~paste(Country,"Volume: ",TotalValue), fillOpacity = 0.5
  ) %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "Brazil"], countries$latitude[countries$Country == "Brazil"], color = 'green',
           flow = USA_group_joined[USA_group_joined[,Year == 2010 & Country == "Brazil"]]$TotalValue, opacity = 0.5) %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "China"], countries$latitude[countries$Country == "China"], color = 'red',
           flow = USA_group_joined[USA_group_joined[,Year == 2010 & Country == "China"]]$TotalValue, opacity = 0.5) %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "Australia"], countries$latitude[countries$Country == "Australia"], color = 'orange',
           flow = USA_group_joined[USA_group_joined[,Year == 2010 & Country == "Australia"]]$TotalValue, opacity = 0.5)


```

Lets do it with time

```{r}
leaflet(USA_group_joined) %>% addTiles() %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "Brazil"], countries$latitude[countries$Country == "Brazil"], color = 'green',
           flow = USA_group_joined[USA_group_joined[, Country == "Brazil"]]$TotalValue, opacity = 0.5, time = unique(USA_group_joined$Year)) %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "China"], countries$latitude[countries$Country == "China"], color = 'red',
           flow = USA_group_joined[USA_group_joined[, Country == "China"]]$TotalValue, opacity = 0.5, time =unique(USA_group_joined$Year)) %>%
  leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "Australia"], countries$latitude[countries$Country == "Australia"], color = 'orange',
           flow = USA_group_joined[USA_group_joined[, Country == "Australia"]]$TotalValue, opacity = 0.5, time =unique(USA_group_joined$Year)) %>%
    leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],
           countries$longitude[countries$Country == "Albania"], countries$latitude[countries$Country == "Albania"], color = 'black',
           flow = USA_group_joined[USA_group_joined[, Country == "Albania"]]$TotalValue, opacity = 0.5, time =unique(USA_group_joined$Year))
```


```{r}

basemap = leaflet(USA_group_joined) %>% addTiles()

basemap %>% leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],USA_group_joined$longitude, USA_group_joined$latitude,
    flow = USA_group_joined$TotalValue,
    time = USA_group_joined$Year)
```

USA Group Joined cut to show select countries.

```{r}
USA_group_joined_cut = USA_group_joined[Country %in% c("Canada", "China", "Brazil", "India", "Australia", "South Africa", "Mexico")]
```

Repeat the plot

```{r}
basemap = leaflet(USA_group_joined_cut) %>% addTiles()

basemap %>% leaflet.minicharts::addFlows(countries$longitude[countries$Country == "United States"], countries$latitude[countries$Country == "United States"],USA_group_joined_cut$longitude, USA_group_joined_cut$latitude,
    flow = USA_group_joined_cut$TotalValue,
    time = USA_group_joined_cut$Year)
```

